Validation of Visual Statistical Inference, Applied to Linear Models

### 2012-07-01 paper_jasa.tex
model for experiment 3 updated
all figures and tables updated after removing duplicate reponse on a lineup

\green{Something probably needs to be said about subjects who do multiple tests, what checks are done.}

@@

\begin{table}[hbtp]
\caption{Number of subjects, gender and total lineups seen, and number of distinct lineups for all three experimental data sets. Note that in some of the lines the number of male and female participants does not add up to the total number of participants due to missing demographic information. }
\begin{center}
\begin{tabular}{ccrrcc}
   \hline 
 Experiment & Subject & Male & Female & Responses &Lineup\\ 
    \hline
  1 & 239 & 121 & 107 & 2317 & 60 \\ 
  2 & 351 & 185 & 164 & 3858 & 70 \\ 
  3 & 156 & 106 &  50 & 1681 & 30 \\ 
   \hline
\end{tabular}
\end{center}
\label{tbl:summary}
\end{table}

@@

\begin{table}[hbtp]
\caption{Parameter estimates of model in Equation \ref{eqn:mixed}. Estimates are highly significant with $p$-value $<$  0.0001 for all three experiment data.}
\begin{center}
\begin{tabular}{cccc}
  \hline
 &  \multicolumn{2}{c} {Fixed effect}  & Random effect\\
 \cline{2-3}
 Experiment & Estimate & Std. error & Variance\\
  \hline
  1 & 0.39 & 0.0093 & 0.0077 \\ 
  2 & 1.22 &  0.0196 &  0.0455 \\ 
  3 &  & & \\ 
   \hline
\end{tabular}
\end{center}
\label{tbl:model_par}
\end{table}

@@

\green{The dashed line cannot be seen in experiment 2 for small effect, and it is essential to see this at this point. What can you do to make the dashed line visible?}

@@

Note that $\hat{p}_D$ is bounded within $(0, 1/(m-1))$. This estimate has to be interpreted carefully, since the choice of a large lineup size, $m$, will make the value look `significant' in a traditional sense. Values close to the upper bound of $1/(m-1)$ should therefore be interpreted as `at least as high', rather than their actual numeric value.
%\hh{For $p$-values below $1/(m-1)$ visual $p$-values are very similar to actual $p$-values, the correlation for those values is: 0.48, 0.85, and 0.11 based on $n_1 = 39, n_2 = 41$, and $n_3= 18$ lineups, respectively.  }  \blue{This is not happening as we see in figure \ref{fig:pval_plot_signal}. But it is true for experiment 2, right?. }
%is equivalent to classical test $p$-value when $p$-value is less than or equal to $1/(m-1)$.  
%and beyond that the plot signal strength levels off on a point bigger than 0.05. Thus if plot signal strength is less than 0.05, we may decide to reject the null hypothesis.

>>>>

\green{Something probably needs to be said about subjects who do multiple tests, what checks are done.} \blue{We do have some data where we notice multiple responses for a lineup by a subject. This happens when subject accidentally use browser's ``back'' button and same data is saved twice. We made sure in the data cleaning process that we do not include more than one response on a lineup by a subject. For this we only keep the very first response.}

@@

\begin{table}[hbtp]
\caption{Number of subjects, gender and total lineups seen, and number of distinct lineups for all three experimental data sets. Note that in some of the lines the number of male and female participants does not add up to the total number of participants due to missing demographic information. \blue{Notice that we do not have data on one lineup (so we have only 29 lineups) for experiment 3 due to the screening process.}}
\begin{center}
\begin{tabular}{ccrrcc}
   \hline 
 Experiment & Subject & Male & Female & Responses &Lineup\\ 
    \hline
1 & 239 & 121 & 107 & 2249 &  60 \\ 
  2 & 351 & 185 & 164 & 3636 &  70 \\ 
  3 & 155 & 103 &  52 & 1511 &  29 \\  
   \hline
\end{tabular}
\end{center}
\label{tbl:summary}
\end{table}

@@

\begin{table}[hbtp]
\caption{Parameter estimates of model in Equation \ref{eqn:mixed}. Estimates are highly significant with $p$-value $<$  0.0001 for all three experiment data.}
\begin{center}
\begin{tabular}{cr@{.}lcc}
  \hline
 &  \multicolumn{3}{c} {Fixed effect}  & Random effect\\
 \cline{2-4}
 Experiment & \multicolumn{2}{l}{Estimate}  &Std. error & Variance\\
  \hline
  1 & 0&39 & 0.0094 & 0.0080 \\ 
  2 & 1&21 &  0.0197 &  0.0443 \\ 
  3 & 0&31 (Intercept)  &   0.1110 & 0.1407\\ 
     & 0&34 (Slope)    &  0.0581     & \\ 
     &-0&72 (correlation) & & \\
   \hline
\end{tabular}
\end{center}
\label{tbl:model_par}
\end{table}

@@

\green{The dashed line cannot be seen in experiment 2 for small effect, and it is essential to see this at this point. What can you do to make the dashed line visible?} \blue{The dashed line layer is placed on top of all other layers. Still there is trouble to see it. Do you think it will work?}

@@

Note that $\hat{p}_D$ is bounded within $(0, 1/(m-1))$. This estimate has to be interpreted carefully, since the choice of a large lineup size, $m$, will make the value look `significant' in a traditional sense. Values close to the upper bound of $1/(m-1)$ should therefore be interpreted as `at least as high', rather than their actual numeric value.
%\hh{For $p$-values below $1/(m-1)$ visual $p$-values are very similar to actual $p$-values, the correlation for those values is: 0.36, 0.79, and -0.48 based on $n_1 = 44, n_2 = 45$, and $n_3= 26$ lineups, respectively.  }  \blue{This is not happening as we see in figure \ref{fig:pval_plot_signal}. But it is true for experiment 2, right?. }
%is equivalent to classical test $p$-value when $p$-value is less than or equal to $1/(m-1)$.  
%and beyond that the plot signal strength levels off on a point bigger than 0.05. Thus if plot signal strength is less than 0.05, we may decide to reject the null hypothesis.
